---
title: "Research Outputs in Luxembourg"
format:
  typst:
    toc: true
    section-numbering: 1.1.a
---

```{r}
#| include: false
library(targets)
library(tinytable)

tar_load(type_doi_missing)
tar_load(lu_first_authors)
tar_load(primary_domain_lu)
```

## Introduction

For this study, data from OpenAlex is used to assemble a comprehensive corpus of
scientific works affiliated with Luxembourg. OpenAlex is an open, transparent,
and freely accessible scholarly metadata platform that aggregates detailed
information on publications, authors, institutions, funding, and collaborations
across all academic disciplines.

Its inclusive coverage and up-to-date data make OpenAlex an ideal resource for
evidence-based policy analysis and strategic decision-making in research and
higher education. By leveraging OpenAlex, this study ensures a broad and
reliable overview of Luxembourg’s research output, enabling the Ministry of
Research and Higher Education to accurately assess national strengths,
collaboration patterns, and areas for future investment.

However, it is important to note that OpenAlex’s coverage, while extensive, is
not yet fully comprehensive for Luxembourg. Some publications, especially those
from smaller institutions, less internationally visible disciplines, or
published in languages other than English may be underrepresented. Given the
complexity of scholarly publishing and data aggregation, it is difficult to
precisely identify all gaps or biases in the data. This uncertainty should be
considered when interpreting the findings and formulating policy
recommendations.

## Exploring the Data: Understanding Luxembourg’s Research Output in OpenAlex

The scope of this study are the scholarly works from the last 10 years. The
following table shows the distribution of type of works and whether they have a
DOI or not:

```{r}
#| echo: false
#| label: tbl-type-doi-missing
#| tbl-cap: Distribution of types of work and missingness of DOI

tt(type_doi_missing)
```

Because the majority of works are articles, this study is going to focus on
articles only, regardless of wheter they have a DOI or not.

```{r}
#| echo: false
#| label: tbl-lu-first-authors
#| tbl-cap: Number of articles where the first author is LU-affiliated

tt(lu_first_authors)
```


```{r}
#| echo: false
#| label: tbl-primary-domain-lu
#| tbl-cap: Number of articles where the first author is LU-affiliated

tt(primary_domain_lu, caption = "Distribution of primary domains by affiliation")
```



To provide a solid foundation for analysis, it is crucial first to understand
the scope and composition of the dataset extracted from OpenAlex. This section
presents key descriptive statistics that characterize Luxembourg’s scientific
output captured in the data.

Total Number of Works The dataset includes all documented research works
affiliated with Luxembourgish institutions or authors over the study period.
This total volume provides a baseline for assessing the country’s overall
scientific productivity.

Works by Luxembourg Institutions Breaking down the dataset by institution allows
us to identify the main contributors within Luxembourg’s research ecosystem.
This includes major universities, research institutes, and other organizations
based in the country, highlighting their relative publication activity.

Works with Primary Luxembourgish Authors In addition to institutional
affiliation, the nationality or primary affiliation of the lead authors provides
insight into the domestic research capacity. We examine the number and share of
works where the primary or corresponding author is Luxembourgish, which helps
distinguish between research led locally versus research involving Luxembourg
through collaboration.

These basic metrics help frame further analysis, such as collaboration patterns,
disciplinary strengths, and research impact, by clarifying what the data
contains and its coverage limits.
